First, I offer my gratitude to my supervisor, Dr.Marin Litoiu whose support throughout my Ph.D. and his assistance especially during the final years of my Ph.D. studies has been vital. I attribute the level of my Ph.D. degree to his encouragement and effort; without him, this thesis not have been completed.

I warmly thank other members of Center of Excellence for Research in Adaptive Systems. Special thanks to my fellow Ph.D. student Cornel Barna for his constructive comments during my Ph.D., to Damon Sotoudeh for being my best schoolmate, and to Kent Poots for providing lots of positive feedbacks on the structure of the thesis. 

Very special thanks to my greatest computer science teachers who truly made a difference in my life, Dr. Amir Amintabar, and Professor Mohammad Taghi Rouhani Rankouhi. I should also thank Prof. Stephen P. Boyd and Stanford University for providing several freely accessible online courses in the area of optimization, as well as their open-source software CVX.  

I must also acknowledge Mohan Mishra, Sheila Wilmot and other York University graduate students union staff, for their critical assistance and help. 

I owe my loving thanks to my parents, my mother Nahid and my father Behrooz, for supporting me throughout all my studies. I should also thank my mother and father-in-law Mahvash and Malek Niaz for their kindness and continued support.  

Finally, and most importantly, I would like to thank my wife Elham for sharing six years of her precious life with me; for tolerating all of my shortcomings, and continuously injecting positive energy into our lives.  



