% \section{Co-Authorship}
This thesis is based my published work listed below. In these publications, I am the main author and the major contributor of the research work presented. In particular, I contributed in the following ways except where noted: materializing the initial idea, researching related work, conducting experiments and analyzing the resulting data (developing required tools). 

Parts of this thesis have been published as follows:
% \newcounter{saveenum} 
\begin{enumerate}
%\setcounter{enumi}{\value{saveenum}}
\item \textbf{Hamoun Ghanbari}, Marin Litoiu, Cornel Barna, Przemyslaw Pawluk, Replica placement in cloud through simple stochastic model predictive control, in 7th IEEE international conference on cloud computing, Alaska, USA, 2014. 
\item \textbf{H. Ghanbari}, B. Simmons, M. Litoiu, C. Barna, and G. Iszlai, Optimal autoscaling in a IaaS cloud, in Proceedings of the 9th international conference on Autonomic computing, San Jose, California, USA, 2012, pp. 173-178.
\item \textbf{H. Ghanbari}, B. Simmons, M. Litoiu, and G. Iszlai, Feedback-based optimization of a private cloud, Future Generation Computer Systems, vol. 28, no. 1, pp. 104-111, Jan, 2012.
\item \textbf{Hamoun Ghanbari}, Bradley Simmons, Marin Litoiu, and Gabriel Iszlai, Exploring alternative approaches to implement an elasticity policy, in Proceedings of the 4th IEEE international conference on cloud computing, Washington DC, USA, 2011, pp. 716 - 723. 
\item \textbf{H. Ghanbari et al.}, Tracking adaptive performance models using dynamic clustering of user classes, in Proceeding of the second joint WOSP/SIPEW international conference on performance engineering, San Jose, California, USA, 2010, pp. 179-188.
\item \textbf{Hamoun Ghanbari} and Marin Litoiu, Identifying implicitly declared self-tuning behavior through dynamic analysis, in ICSE workshop on software engineering for adaptive and self-managing systems, Vancouver, BC, Canada, 2009, pp. 48-55.
\item \textbf{Hamoun Ghanbari}, Autonomic mechanisms in cloud computing ecosystems, Qualifying exam report, 2011, \url{http://www.cse.yorku.ca/~hamoun/qual/}. 
\end{enumerate}
